---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# treelabel

<!-- badges: start -->
<!-- badges: end -->

The goal of treelabel is to allow storing labels which exist in a hierarchical relationship.

## Installation

You can install the development version of treelabel like so:

``` r
devtools::install_github("const-ae/treelabel")
```

## Example

We begin by defining our label hierarchy using [`igraph`](https://r.igraph.org/articles/igraph.html).

```{r}
tree <- igraph::graph_from_literal(
  Animal - Bird : Mammal,
  Mammal - Cat : Dog,
  Dog - Retriever : Poodle,
  Bird - Parrot : Eagle,
  Parrot - Ara  : Budgie
)
plot(tree, layout = igraph::layout_as_tree(tree, root = "Animal"),
     vertex.size = 40, vertex.label.cex = 0.8)
```


If you provide a character vector and a tree defining the hierarchy, you get a `treelabel` vector that is backed by a matrix of `TRUE` / `FALSE` values indicating the logical relationships.

```{r}
library(treelabel)
vec <- treelabel(c("Ara", "Eagle", "Cat", NA, "Ara", "Poodle", "Cat", "Cat"), tree = tree, tree_root = "Animal")
vec
```

The underlying matrix:
```{r}
tl_score_matrix(vec)
```


You can test if an element is of a specific type with the `tl_get` function.

```{r}
# Is the element an eagle?
tl_get(vec, "Eagle")

# Is the element a bird?
tl_get(vec, "Bird")
```


You can also express uncertainty about any particular label

```{r}
vec2 <- treelabel(list(
  c(Dog = 1, Poodle = 0.8, Retriever = 0.2),
  c(Cat = 0.9),
  c(Animal = 0.78)
), tree, tree_root = "Animal")

treelabel(tree = tree, tree_root = "Animal")

vec2
```

Concatenation

```{r}
c(tl_as_numeric(vec), vec2)

# c("Eagle", vec)
c(vec, "Eagle")
```



```{r, paged.print=FALSE}
library(tidyverse)
tibble(id = seq_len(8), vec) |>
  filter(tl_eval(vec, Bird))
```


# Cell type scores

```{r}
edges <- c("root", "immune cell",
           "root", "epithelial cell",
           "immune cell", "myeloid cell",
           "immune cell", "lymphoid cell",
           "lymphoid cell", "t cell",
           "lymphoid cell", "b cell",
           "t cell", "CD4 t cell",
           "t cell", "CD8 t cell",
           "t cell", "treg cell",
           "myeloid cell", "neutrophil",
           "myeloid cell", "dendritic cell",
           "myeloid cell", "macrophage")

g <- igraph::graph(edges, directed = TRUE)
plot(g, layout = igraph::layout_as_tree(g, root = "root"))
```


```{r}
label_list <- list(c("immune cell" = 1, "lymphoid cell" = 0.99, "b cell" = 0.7),
     c("immune cell" = 1, "lymphoid cell" = 0.99, "t cell" = 0.6, "b cell" = 0.3, "CD4 t cell" = 0.59),
     c("immune cell" = 1, "lymphoid cell" = 0.99, "t cell" = 0.6, "b cell" = 0.3, "CD4 t cell" = 0.59),
     c("immune cell" = 1, "lymphoid cell" = 0.99, "t cell" = 0.6, "b cell" = 0.2, "CD8 t cell" = 0.59),
     c("immune cell" = 1, "lymphoid cell" = 0.99, "t cell" = 0.89, "b cell" = 0.1, "CD4 t cell" = 0.89),
     c("myeloid cell" = 1),
     c("myeloid cell" = 1, "dendritic cell" = 1))
labels1 <- treelabel(label_list, tree = g)

labels1
```


Arithmetics

```{r}
mean_lab <- mean(labels1, na.rm=TRUE)
mean_lab
tl_score_matrix(mean_lab)
```


```{r, paged.print=FALSE}
tibble(id = seq_along(labels1), label1 = labels1) |>
  mutate(new_score = tl_eval(label1, log(`t cell` + 17)))
```


```{r, paged.print=FALSE}
labels2 <- treelabel(c("t cell", "CD4 t cell", "CD4 t cell", "dendritic cell", "CD4 t cell", "dendritic cell", "dendritic cell"), g, "root")

df <- tibble(id = seq_along(labels1), 
             label1 = labels1, 
             label2 = labels2)
df

df |>
  mutate(new = tl_mean_across(starts_with("label"), `CD4 t cell` > 0.2))

df |>
  filter(tl_if_all(starts_with("label"), `t cell` > 0.2))

df |>
  mutate(n_tcell = tl_count_across(starts_with("label"), `CD4 t cell` > 0.2),
         frac_tcell = tl_fraction_across(starts_with("label"), `CD4 t cell` > 0.2),
         mean_tcell = tl_mean_across(starts_with("label"), `CD4 t cell` > 0.2))

```


```{r}
label_list <- list(c("immune cell" = 1, "lymphoid cell" = 0.99, "b cell" = 0.7, "t cell" = 0.12),
                   c("immune cell" = 1, "lymphoid cell" = 0.99, "b cell" = NA, "t cell" = 0.99, "CD4 t cell" = 0.8))
lab3 <- treelabel(label_list, tree = g)
lab3

tl_eval(lab3, `CD4 t cell` > 0.6)
tl_eval(lab3, `CD4 t cell` > 0.6) & tl_eval(lab3, `t cell` > 0.6)
pmin(tl_eval(lab3, `CD4 t cell` > 0.6), tl_eval(lab3, `t cell` > 0.6), na.rm=TRUE)
tl_eval(lab3, `t cell`)

tl_score_matrix(lab3)
tl_is(lab3, `CD4 t cell`)
tl_is(labels1, `t cell` > 0.2)
```



# Session Info

```{r}
sessionInfo()
```


